---
title: "Data_cleaning"
author: "Felix Tran"
date: "November 27, 2018"
output: html_document
---

This Rmarkdown file contains code for reading in and cleaning the raw dataset.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

Read in dataset. Looking at the variable for year, some observations had values
which do not make sense. We thus kept only the observations with values for
the variable **year_description** that make sense. Cleaned variable names.
```{r}
asthma_df <- readr::read_csv('./Data/Air_Quality.csv')

asthma_df %>% 
  group_by(year_description) %>% 
  summarize(counts = n())

asthma_df <- asthma_df %>% 
  filter(year_description %in% c('2005', '2005-2007', '2009-2011', '2013')) %>% 
  janitor::clean_names()
```

Looking at the different observations recorded in the dataset. There are 18
unique kinds of observations recorded for NYC overall, by boroughs, and by
neighborhoods for the corresponding years. 
```{r}
asthma_df %>% 
  mutate(obsv = stringr::str_c(name, measure, sep = " ")) %>% 
  group_by(obsv) %>% 
  summarize(counts = n())
```

Save this cleaned dataset in a csv file. 
```{r}
readr::write_csv(asthma_df, path = './Data/nyc_asthma_cleaned.csv')
```