---
title: "Linear regression"
author: "Felix Tran"
date: "December 2, 2018"
output: html_document
---

This file runs a linear regression on PM 2.5-attributable ED visits among NYC
neighborhoods by PM 2.5 levels and neighborhood poverty. Neighborhoods are 
defined by the [United Hospital Fund 42 neighborhood designations](https://www.health.ny.gov/statistics/cancer/registry/appendix/neighborhoods.htm)

<br><br>

# The regression model

$Asthma\:ED\:rate = Intercept + PM\:2.5\:levels + Neighborhood\:poverty$

<br><br>

# The data

### Rate of PM 2.5-attributable ED visits per 100,000
Rates of ED visits were obtained from NYC Open Data for the timeframe 2009-2011. 
Rates were reported for children 0-17 years old and adults 18 years and older by
neighborhood. 


### PM 2.5 air pollution
PM 2.5 air levels are measured as average micrograms of PM 2.5 per cubic meter 
of air for a given year. Measures of PM 2.5 air pollution were obtained from the 
NYC Environment and Health Data Portal. Measures cover the timeframe 2009-2011.


### Neighborhood poverty
Measure of neighborhood poverty were obtained from the NYC Environment and 
Health Data Portal. Measure cover the timeframe 2009-2013. This is the smallest
timeframe in which a neighborhood-level poverty rate can be estimated. Although
this timeframe does not exactly match the timeframe of the other variables, we
believe this estimate is not too far off from the true rate of poverty in NYC
neighborhoods for 2009-2011. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(modelr)
library(mgcv)
library(purrr)
library(readr)
```

# Cleaning the data

First, the data are cleaned and merged by neighborhood using a unique ID
assigned to each neighborhood.
```{r}
poverty_df <- readr::read_csv('./Data/pm_and_poverty_09-13.csv') %>% 
  janitor::clean_names() %>% 
  filter(str_detect(geo_type_name, "UHF") == T) %>% 
  filter(measure == "Percent") %>%
  filter(name == "Poverty") %>% 
  mutate(geography_id = as.integer(geo_entity_id)) %>% 
  rename(poverty_rate = data_value) %>% 
  select(geography_id, poverty_rate)

pm_df <- readr::read_csv('./Data/pm_and_poverty_09-13.csv') %>% 
  janitor::clean_names() %>% 
  filter(str_detect(geo_type_name, "UHF") == T) %>% 
  filter(measure == "Mean") %>%
  filter(str_detect(name, "PM") == T) %>% 
  mutate(geography_id = as.integer(geo_entity_id)) %>% 
  group_by(geography_id) %>% 
  summarize(pm_level = round(mean(data_value), digits = 3))

asthma_df <- readr::read_csv('./Data/nyc_asthma_cleaned.csv') %>% 
  janitor::clean_names() %>% 
  filter(str_detect(geo_type_name, "UHF") == T) %>% 
  filter(str_detect(name, "PM2.5-Attributable Asthma ED") == T) %>% 
  filter(year_description == '2009-2011') %>% 
  rename(geography_id = geo_entity_id) %>% 
  distinct() %>% 
  inner_join(., poverty_df, by = "geography_id") %>% 
  inner_join(., pm_df, by = "geography_id") %>% 
  select(-c(indicator_data_id, indicator_id, name, geo_type_name, 
            year_description)) %>% 
  spread(key = measure, value = data_valuemessage) %>% 
  janitor::clean_names() %>% 
  rename(child_rate = rate_children_0_to_17_yrs_old,
         adult_rate = rate_18_yrs_and_older)
```

<br><br>

# Regression on adult ED visits
```{r, echo = F}
calc_lower_ci <- function(beta, stnd.error) {
  lower = round(beta - (1.96 * stnd.error), digits = 2)
  lower
}

calc_upper_ci <- function(beta, stnd.error) {
  lower = round(beta + (1.96 * stnd.error), digits = 2)
  lower
}

adult_reg_df <- lm(data = asthma_df, adult_rate ~ poverty_rate + pm_level) %>% 
  broom::tidy() %>% 
  mutate(upper_ci = calc_upper_ci(estimate, std.error),
         lower_ci = calc_lower_ci(estimate, std.error),
         ci = str_c(lower_ci, upper_ci, sep = " to "),
         p.value = round(p.value, digits = 3)) %>% 
  select(term, estimate, ci, p.value)

adult_reg_df %>% 
  knitr::kable(col.names = c("Term", "Beta coefficient", 
                             "95% CI", "P-value"),
               format = "html",
               caption = "Rates of PM 2.5-attributable adult ED visits among NYC
               neighborhoods in 2009-2011 by PM 2.5 levels and 
               neighborhood poverty",
               align = 'c')
```
<br><br>
$Asthma\:ED\:Rate = -42.67 + 3.46(Neighborhood\:poverty) + 2.02(PM\:2.5\:air\:levels)$

For every 1 unit increase in PM 2.5 air levels, adult PM 2.5-attributable asthma
ED visits per 100,000 adults for a given neighborhood increase on average by 
2.02 visits. The confidence interval (-3.9 to 7.93) suggests this observed 
association may be spurious. 

For every 1 percentage point increase in the proportion of people living in 
poverty in a neighborhood, adult PM 2.5-attributable asthma ED visits per 
100,000 adults for a given neighborhood increase on average by 3.46 visits. The
confidence interval (2.71 to 4.2) suggests that we are unlikely to observe
this association if neighborhood poverty truly has no association with asthma ED 
rates.

<br><br>

# Regression on child ED visits
```{r, echo = F}
child_reg_df <- lm(data = asthma_df, child_rate ~ poverty_rate + pm_level) %>% 
  broom::tidy() %>% 
  mutate(upper_ci = calc_upper_ci(estimate, std.error),
         lower_ci = calc_lower_ci(estimate, std.error),
         ci = str_c(lower_ci, upper_ci, sep = " to "),
         p.value = round(p.value, digits = 3)) %>% 
  select(term, estimate, ci, p.value)

child_reg_df %>% 
  knitr::kable(col.names = c("Term", "Beta coefficient", 
                             "95% CI", "P-value"),
               format = "html",
               caption = "Rates of PM 2.5-attributable child ED visits among NYC
               neighborhoods in 2009-2011 by PM 2.5 levels and 
               neighborhood poverty",
               align = 'c')
```
<br><br>
$Asthma\:ED\:Rate = -99.61 + 5.66(Neighborhood\:poverty) + 8.29(PM\:2.5\:air\:levels)$

For every 1 unit increase in PM 2.5 air levels, child PM 2.5-attributable asthma
ED visits per 100,000 children for a given neighborhood increase on average by 
8.29 visits. The 95% confidence interval (-3.96 to 20.53) suggests this observed 
association may be spurious. 

For every 1 percentage point increase in the proportion of people living in 
poverty in a neighborhood, child PM 2.5-attributable asthma ED visits per 
100,000 children for a given neighborhood increase on average by 5.66 visits. The
95% confidence interval (4.12 to 7.2) suggests that we are unlikely to observe
this association if neighborhood poverty truly has no association with asthma ED 
rates.

<br><br>

# Comments
For both child and adult rates of ED vists, levels of neighborhood PM 2.5 air
pollution do not appear to have a statistically significant effect. Rather, 
neighborhood poverty is more strongly associated with asthma ED visits.
